---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(lgr)
# <img src="man/figures/lgr-logo-plain.svg" align="right" width=160 height=160/>
```
# lgr

[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/s-fleck/lgr.svg?branch=master)](https://travis-ci.org/s-fleck/lgr)
[![Codecov test coverage](https://codecov.io/gh/s-fleck/lgr/branch/master/graph/badge.svg)](https://codecov.io/gh/s-fleck/lgr?branch=master)

lgr is a logging package for R built on the back of 
[R6](https://github.com/r-lib/R6) classes. It is designed to be flexible,
performant and extensible. 

Users that have not worked with R6 classes before, will find the syntax for
configuring Loggers a bit strange, but I did my best to compose a 
hopefully helpful [vignette](https://s-fleck.github.io/lgr/articles/lgr.html). 
Users that come from python or Java, will feel at home as lgr borrows heavily 
from [Apache Log4j](https://logging.apache.org/log4j/2.x/) and
[Python logging](https://docs.python.org/3/library/logging.html).


## Features

* *Hierarchical loggers* like in log4j and python logging. This is useful if you
  want to be able to configure logging on a per-package basis.
* An *arbitrary number of appenders* for each logger. A single logger can write
  values to the console, a logfile, a database, etc... .
* Allow for *custom fields* in log events. As opposed to many other logging 
  packages for R a log event is not just a message with a timestamp, but
  can contain arbitrary data fields. This is very helpful if you want to produce
  logs that are machine readable and easy to analyze.
* *Vectorized* logging (so `lgr$fatal(capture.output(iris))` works)
* Lightning fast in-memory log based in `data.table` included for 
  interactive use.
* Support for advanced appenders such as database appenders, cached appenders,
  etc... .
* Optional *color* support via [crayon](https://github.com/r-lib/crayon)





## Usage

To log an *event* with with lgr we call `lgr$<logging function>()`. Unnamed 
arguments to the logging function are interpreted by `sprintf()`.

```{r}
lgr$fatal("A critical error")
lgr$error("A less severe error")
lgr$warn("A potentially bad situation")
lgr$info("iris has %s rows", nrow(iris))

# the following log levels are hidden by default
lgr$debug("A debug message")
lgr$trace("A finer grained debug message")
```

A Logger can have several Appenders. For example, we can add a JSON appender to
log to a file with little effort. 

```{r}
tf <- tempfile()
lgr$add_appender(AppenderJson$new(tf))
lgr$info("cars has %s rows", nrow(cars))

cat(readLines(tf))
```

JSON naturally supports custom fields. Named arguments passed to `info()`, 
`warn()`, etc... are intepreted as custom fields.

```{r}
lgr$info("loading cars", "cars", rows = nrow(cars), cols = ncol(cars))

cat(readLines(tf), sep = "\n")

```

```{r, echo=FALSE}
unlink(tf)

```

For more examples please see the 
package 
[vignette](https://s-fleck.github.io/lgr/articles/lgr.html) and 
[documentation](https://s-fleck.github.io/lgr/)





## Development Status

lgr is stable, but I want to conduct a some field-tests to iron out the last
few bugs before I announce it. A CRAN release is planned for February 2019.




## Dependencies

[R6](https://github.com/r-lib/R6): The R6 class system prevents the framework
on which lgr is built and the **only Package lgr will ever depend on**.

### Optional Dependencies

These optional dependencies that are not necessary to use lgr, but
that are required for some extra appenders. Care was taken to choose 
packages that are slim, stable, have minimal dependencies, and are well 
mentained:

  * [crayon](https://github.com/r-lib/crayon) for colored console output.
  * [data.table](https://github.com/Rdatatable/) for fast in-memroy logging
    with `AppenderDt`. 
  * [jsonlite](https://github.com/jeroen/jsonlite) for JSON logging via 
    `LayoutJson`. JSON is a populat plaintext based file format that is easy to 
    read for humans and machines alike.
  * [DBI](https://github.com/r-dbi/DBI) for logging to databases. Logging with
    lgr has been tested with the following backends: 
    - [RSQLite](https://github.com/r-dbi/RSQLite), 
    - [RMariaDB](https://cran.r-project.org/web/packages/RMySQL/index.html) for MariaDB and MySQL,
    - [RPostgreSQL](https://cran.r-project.org/web/packages/RPostgreSQL/index.html),
    - [RJDBC](https://github.com/s-u/RJDBC) for DB2. 
    
    In theory all DBI compliant database packages should work. If you
    are using lgr with a database backend, please report your (positive and
    negative) experiences to me.
  * [whoami](https://github.com/r-lib/whoami/blob/master/DESCRIPTION) for 
    guessing the user name from various sources. You can also set the user name 
    manually if you want to use it for logging.
    

## Installation

``` r
devtools::install.github("s-fleck/lgr")
```

## Outlook

The long term goal is to support (nearly) all features of the python logging
module. If you have experience with python logging or Log4j and are missing
features/appenders that you'd like to see, please feel free to post a feature 
request on the issue tracker.


## Acknowledgements

* [Inkscape](https://inkscape.org/) for the hex sticker
* [draw.io](https://draw.io/) for the flow chart in the vignette
