---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(lgr)
# <img src="man/figures/lgr-logo-plain.svg" align="right" width=160 height=160/>
```
# lgr

[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/s-fleck/lgr.svg?branch=master)](https://travis-ci.org/s-fleck/lgr)
[![Codecov test coverage](https://codecov.io/gh/s-fleck/lgr/branch/master/graph/badge.svg)](https://codecov.io/gh/s-fleck/lgr?branch=master)

lgr is a logging package for R built on the back of 
[R6](https://github.com/r-lib/R6) classes. It is designed to be flexible,
performant and extensible. 

Users that have not worked with R6 classes before, will find the syntax for
configuring Loggers a bit strange, but I did my best to compose a 
hopefully helpful [vignette](https://s-fleck.github.io/lgr/articles/lgr.html). 
User that have experience with 
  [shiny](https://github.com/rstudio/shiny), 
  [plumber](https://github.com/trestletech/plumber),
  [python logging](https://docs.python.org/3/library/logging.html) or 
  [Apache Log4j](https://logging.apache.org/log4j/2.x/) before will fill feel
at home.



## Features

* *Hierarchical loggers* like in log4j and python logging. This is useful if you
  want to be able to configure logging on a per-package basis.
* An *arbitrary number of appenders* for each logger. A single logger can write
  to the console, a logfile, a database, etc... .
* Allow for *custom fields* in log events. As opposed to many other logging 
  packages for R a log event is not just a message with a timestamp, but
  can contain arbitrary data fields. This is very helpful if you want to produce
  logs that are machine readable and easy to analyze.
* *Vectorized* logging (so `lgr$fatal(capture.output(iris))` works)
* Lightning fast *in-memory log* based in `data.table` included for 
  interactive use.
* Comes with a *wide range of appenders*, for example for:
    * Appending to Databases (buffered or directly)
    * Sending notifications via email or pushbullet
    * writing JSON with arbitrary data fields
    * In memory buffers
    * colored console output
* Optional support to use [glue](https://glue.tidyverse.org/) instead of 
  `sprintf()` for composing log messages.


## Usage

To log an *event* with with lgr we call `lgr$<logging function>()`. Unnamed 
arguments to the logging function are interpreted by `sprintf()`. For a way
to create loggers that [glue](https://glue.tidyverse.org/) instead
please refer to the vignette.

```{r}
lgr$fatal("A critical error")
lgr$error("A less severe error")
lgr$warn("A potentially bad situation")
lgr$info("iris has %s rows", nrow(iris))

# the following log levels are hidden by default
lgr$debug("A debug message")
lgr$trace("A finer grained debug message")
```

A Logger can have several Appenders. For example, we can add a JSON appender to
log to a file with little effort. 

```{r}
tf <- tempfile()
lgr$add_appender(AppenderJson$new(tf))
lgr$info("cars has %s rows", nrow(cars))
cat(readLines(tf))
```

JSON naturally supports custom fields. Named arguments passed to `info()`, 
`warn()`, etc... are interpreted as custom fields.

```{r}
lgr$info("loading cars", "cars", rows = nrow(cars), cols = ncol(cars))
cat(readLines(tf), sep = "\n")
```

```{r, echo=FALSE}
unlink(tf)
```

For more examples please see the 
package 
[vignette](https://s-fleck.github.io/lgr/articles/lgr.html) and 
[documentation](https://s-fleck.github.io/lgr/)




## Development Status

The api of lgr is stable and safe for use. The internal implementation of
the database logging features still needs some refinement, and if you are
using lgr with a database, I would be grateful for any kind of feedback.




## Dependencies

[R6](https://github.com/r-lib/R6): The R6 class system prevents the framework
on which lgr is built and the **only Package lgr will ever depend on**.

### Optional Dependencies

These optional dependencies that are not necessary to use lgr, but
that are required for some extra appenders. To leverage the power of the most
central features of **lgr** it is recommended (but not necessary) that you
at least install the packages **data.table**, **jsonlite**, 
and possibly **crayon**. Care was taken to choose packages that are slim, 
stable, have minimal dependencies, and are well 
maintained :

  * [crayon](https://github.com/r-lib/crayon) for colored console output.
  * [data.table](https://github.com/Rdatatable/) for fast in-memory logging
    with `AppenderDt`. 
  * [jsonlite](https://github.com/jeroen/jsonlite) for JSON logging via 
    `LayoutJson`. JSON is a popular plaintext based file format that is easy to 
    read for humans and machines alike.
  * [DBI](https://github.com/r-dbi/DBI) for logging to databases. Logging with
    lgr has been tested with the following backends: 
    - [RSQLite](https://github.com/r-dbi/RSQLite), 
    - [RMySQL](https://cran.r-project.org/package=RMySQL) for MariaDB and MySQL (RMariaDB is currently broken, see [this issue](https://github.com/r-dbi/RMariaDB/issues/119)),
    - [RPostgreSQL](https://cran.r-project.org/package=RPostgreSQL),
    - [RJDBC](https://github.com/s-u/RJDBC) for DB2. 
    
    In theory all DBI compliant database packages should work. If you
    are using lgr with a database backend, please report your (positive and
    negative) experiences, as database support is still somewhat experimental.
  * [gmailr](https://cran.r-project.org/package=gmailr) or
    [sendmailR](https://cran.r-project.org/package=sendmailR) for email 
    notifications.
  * [RPushbullet](https://github.com/eddelbuettel/rpushbullet) for push
    notifications.
  * [whoami](https://github.com/r-lib/whoami/blob/master/DESCRIPTION) for 
    guessing the user name from various sources. You can also set the user name 
    manually if you want to use it for logging.
    

## Installation

``` r
devtools::install.github("s-fleck/lgr")
```

## Outlook

The long term goal is to support (nearly) all features of the python logging
module. If you have experience with python logging or Log4j and are missing
features/appenders that you'd like to see, please feel free to post a feature 
request on the issue tracker.




## Acknowledgement

* [Inkscape](https://inkscape.org/) for the hex sticker
* [draw.io](https://draw.io/) for the flow chart in the vignette
