---
title: "lgr-simple"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lgr-simple}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lgr)
```


### Simplified Logging Syntax

If you care about terse syntax you can use the following `UPPERCASE`^[
If you are wondering why I decided on `ALL-CAPS` for the logging functions: It
distinguishes log calls well from normal code, and also all the good names are 
already taken by base R functions (`log()`, `trace()`, `debug()`). Many R 
packages deal with this issue by using prefixes (like `log_debug()`), but I do 
not like that idea. In any case that is pretty to close to what you get when 
you use the R6 logging syntax] commands to log:

```{r}
FATAL("This is an important message about %s going wrong", "something")
ERROR("A less severe error")
WARN("something likely went wrong")
INFO("Everything is ok")
DEBUG("Debug messages are hidden by default")
console_threshold("debug")  # you must use lower case names here
DEBUG("unless we lower the threshold")
TRACE("Trace is even lower than debug")
```

If you have the package **data.table** installed, lgr does not only log to the 
console, but also to an in memory `data.table`. You can use `show_log()` to 
quickly display this in memory log^[If you want to know more about interacting 
with the data.table Appender see the help for `?AppenderDt`.]. 

```{r}
show_log(3)  # show the last 3 log entries
```
If you looked carefully, you see that both debug messages were logged to 
memory (not only the one that you saw in the console). This is explained in the
section on [thresholds](#thresholds).

