---
title: "Example Configurations for yog"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Adding a Custom Logger to a Package

If you are a package author, it is good practice to define a separate logger
for your package. This gives users the ability to easily enable/disable 
logging on a per-package basis. To do this you must put the following code
into any `.R` file inside the `R/` directory of your package:

```{r}
# mypackage/R/mypackage-package.R
.onLoad <- function(...){
  assign(
    "logger",
    yog::Logger$new(name = "mypackage"),  # the logger name should be the same as the package name
    envir = asNamespace("mypackage")
  )
}

```
After you set this up you can use `logger$fatal()`, `logger$info()`, etc... 
inside your package. You do not have to define any appenders, since all log 
events will get redirected to the Root Logger (see the chapter on hierarchical 
logging in the `yog` main vignette). 



# Recommended Logger Setup

The out-of-the-box setup for the Root Logger is minimal to make it easy
to use for newcommers. I prefer to set up my logger like this:

```{r}
yog$add_appender()
```



