---
title: "Yet Another Log4j Inspired Logging Framework"
author: "Stefan Fleck"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{yog}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(yog)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Yog is a logging framework for R inspired by 
[Apache Log4j](https://logging.apache.org/log4j/2.x/) and
[Python logging](https://docs.python.org/3/library/logging.html).
It is based on [R6 classes](https://github.com/r-lib/R6) that 
focuses on extensibility, ease of use, and performance

Some key features of yog are:

  * use multiple appenders with custom thresholds for a single logger.
  

# Basic usage

## Simple logging

Yog comes with a pre-configured root logger included. If you do not care about 
using multiple loggers, and do not want to deal with R6 classes you still use 
yog as a performant and easy to use logging solution.


```{r}
FATAL("This is an important message about %s going wrong", "something")
ERROR("A less severe error")
WARN("something likely went wrong")
INFO("Everything is ok")
DEBUG("Debug messages are hidden by default")
threshold("debug")  # you must use lower case names here
DEBUG("unless we lower the threshold")
TRACE("Trace is even lower than debug")

```



You can also direclty use Logger R6 objects to log. This package comes with a
default root logger named like the package (`"yog"`). This method of referring
to the logger is more explicit (and therefore arguably better). The examples
in this vignette will moslty use the notation below.

```{r}
yog$fatal("This is an important message about %s going wrong", "->something<-")
yog$trace("Trace messages are still hidden")
yog$threshold <- "trace"
yog$trace("Unless we lower the threshold")
```

This only logs to the console. If you want to redirect the output to a file
you can just add a file appender to yog

```{r}
tf <- tempfile()

# Add a new appender to a logger. We don't have to supply a name, but that
# makes it easier to remove later.
yog$add_appender(AppenderFile$new(file = tf), name = "file")

# configure yog so that it logs everything to the file, but only info and above
# to the console

yog$threshold <- NA
yog$appenders$console$threshold <- "info"
yog$appenders$file$threshold <- NA


yog$info("Another informational message")
yog$debug("A debug message that the console appender doesn't show.")

readLines(tf)

# Remove the appender again
yog$remove_appender("file")
```

## Log Levels

Yog supports the standard log4j log levels outlined bellow. Log levels have
numeric and character representations. Creation of custom log levels are not 
yet supported by yog.

```{r, echo = FALSE}

ll <- data.frame(
  `Level` = seq(100, 600, by = 100),
  `Name` = c("FATAL", "ERROR", "WARNING", "INFO", "DEBUG", "TRACE"),
  `Description` = c(
    "Critical error that leads to program abort. Should always indicate a `stop()` or similar",
    "A severe error that does not trigger program abort",
    "A potentially harmful situation, like `warning()`",
    "An informatinal message on the progress of the application",
    "Finer grained informational messages that are mostly useful for debugging",
    "An even finer grained message than debug"
  )
) 

knitr::kable(ll)

```



# Configuring Loggers


## Structure

* A **Logger** collects information and dispatches it to the appropriate appenders
* An **Appender** writes the log message to destination (the console, a file, 
  a database, etc...)
* A **Layout** is


# Advanced Usage

## Hierarchical Loggers

![](log_flow.svg)

