---
title: "Yet Another Log4j Inspired Logging Framework"
author: "Stefan Fleck"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{yog}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(yog)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Yog is a logging framework for R inspired by 
[Apache Log4j](https://logging.apache.org/log4j/2.x/) and
[Python logging](https://docs.python.org/3/library/logging.html).
It follows a traditional object oriented design implemented through
[R6 classes](https://github.com/r-lib/R6). This enables yog to have a larger
set of features than other logging packages for R, and makes it easy to extend 
for people that have experience with R6 classes.


## Quickstart

Configure a new logger that writes all log messages of level *info* or more to 
a text file and of level *error* or more to a json file.

**Setup:**
```{r}
tfi <- tempfile(fileext = ".info")
tfe <- tempfile(fileext = ".error")
l <- Logger$new(
  "example Logger",
  appenders = list(
    AppenderFile$new(tfi, threshold = "info"),
    AppenderFile$new(tfe, threshold = "error", layout = LayoutJson$new())
  )
)

```

**Usage:**
```{r}
l$info("Everything is okay")
l$error("Everything is NOT okay")
l$fatal("NOTHING is okay")
readLines(tfi)
read_json_lines(tfe)
```

```{r echo=FALSE, results='hide'}
file.remove(tfi, tfe)
```



# Usage

## Structure of the Logging System

If you want custom logging configurations, you have to understand the structure
of the logging process.

* A **Logger** collects information and dispatches it to its *Appenders*, and
  also the *Appenders* of its *Parent Loggers* (also see the section on 
  hierarchical logging)
* An **Appender** writes the log message to destination (the console, a file, 
  a database, etc...).
* A **Layout** is used by an Appender to format LogRecords. For example,
  `AppenderFile` uses `LayoutFormat` by default to write human readable log
  events to a text file, but can also use `LayoutJson` produce machine 
  readable JSON lines logfiles.
* **LogEvents** are produced by the Logger and dispatched to Appenders. They 
  contain all the information that is beeing logged (think of
  it as a row in table). LogEvents usually contain the log level, a timestamp,
  a message, the name of the calling function, and a 
  [reference](https://adv-r.hadley.nz/r6.html#r6-semantics]) to the Logger
  that created it. In addition, a LogEvent can contain any number of custom
  fields, but not all Appenders/Layouts support custom fields. For example
  if you use `AppenderFormat` with `LayoutFormat` you can only use the standard 
  fields in your log message, while `LayoutJson` supports custom fields in 
  a quite natural manner. See [examples 1 & 2](#examples)


### On R6 Classes

The elements described above are R6 classes. R6 is an object orientation
system for R that is used by many popular packages such as shiny, dplyr, 
plumber, roxygen2, and testthat but often behind the scenes and not as exposed
as in yog.

You recognize R6 classes in this package because they are named following the 
UpperCamelCase naming convention. While there is only one kind of Logger and 
one kind of LogEvent, there are several subclasses of Appenders and Layouts.

An introduction to R6 classes is beyond the scope of this document, but you
can find the official documentation [here](https://r6.r-lib.org/) and there is
also this [talk on youtube](https://www.youtube.com/watch?v=3GEFd8rZQgY). In
short R6 classes store data (fields) together with functions (methods) and
have to be instantiated with `<classname>$new()`. So if you want to create
a new `AppenderFile`, you do this by calling 
`AppenderFile$new(file = tempfile())`.



## Log Levels

Yog supports the standard log4j Log Levels outlined bellow. The Log Level of
an event represents its severity. The named log levels are really just 
nicknames vor integer values, and you can use the `character` or `integer` 
representations interchangably. You can also use arbitrary integer values 
(greater than `0`), but you are encouraged to stick to the ones bellow. 

```{r, echo = FALSE}
ll <- data.frame(
  `Level` = c(0, seq(100, 600, by = 100), NA),
  `Name` = c("off", "fatal", "error", "warn", "info", "debug", "trace", "all"),
  `Description` = c(
    "A log level of 0/off tells a Logger or Appender to suspend all logging",
    "Critical error that leads to program abort. Should always indicate a `stop()` or similar",
    "A severe error that does not trigger program abort",
    "A potentially harmful situation, like `warning()`",
    "An informatinal message on the progress of the application",
    "Finer grained informational messages that are mostly useful for debugging",
    "An even finer grained message than debug",
    "A log level of NA/all tells a Logger or Appender to process all log events"
  )
) 

knitr::kable(ll)
```

The list of named log levels is stored as an option 
(`getOption("yog.log_levels")`) and you can use `add_log_levels()` and 
`remove_log_levels()` to define your own named levels if you want to.
There are only predefined logging methods (`yog$fatal()`, etc..) for the
standard log levels though.


## Logging With the Root Logger

Yog comes with a pre-configured root Logger. It is called *root* because
you can [set up a tree of Loggers](#hierarchy) that descent from it, but for 
basic use you will not have to worry about that.


### Simplified Logging Syntax

If you care about terse syntax you can use the following `UPPERCASE`^[
If you are wondering why I decided on `ALL-CAPS` for the logging functions: It
distinquishes log calls well from normal code, and also all the good names are 
already taken by base R functions (`log()`, `trace()`, `debug()`). Many R 
packages deal with this issue by using prefixes (like `log_debug()`), but I do 
not like that idea. In any case that is pretty to close to what you get when 
you use the R6 logging syntax] commands to log:

```{r}
FATAL("This is an important message about %s going wrong", "something")
ERROR("A less severe error")
WARN("something likely went wrong")
INFO("Everything is ok")
DEBUG("Debug messages are hidden by default")
console_threshold("debug")  # you must use lower case names here
DEBUG("unless we lower the threshold")
TRACE("Trace is even lower than debug")
```

If you have the package **data.table** installed, yog does not only log to the 
console, but also to an in memory `data.table`. You can use `show_log()` to 
quickly display this in memory log^[If you want to know more about interacting 
with the data.table Appender see the help for `?AppenderDt`.]. 

```{r}
show_log(3)  # show the last 3 log entries
```
If you looked carefully, you see that both debug messages were logged to 
memory (not only the one that you saw in the console). This is explained in the
section on [thresholds](#thresholds).


### R6 Logging Syntax

You can also directly use Logger R6 objects to log. You can refer to the 
*root* logger with `yog$`. This is especially useful if you are working with
multiple loggers (e.g. if you set up a separate Logger for a package you have
developed). The examples in this vignette mostly use the notation below as it
is more explicit:

```{r}
yog$fatal("This is an important message about %s going wrong", "->something<-")
yog$trace("Trace messages are still hidden")
yog$appenders$console$set_threshold("trace")
yog$trace("Unless we lower the threshold")
yog$appenders$memory$show(3)

```

### Formatting Strings 

You can use `sprintf()` style formatting strings directly in log messages. 
```{r}
yog$info("The sky was the color of %s, tuned to a dead chanel", "television")
INFO("There are %.1f lights", 4.5321)
```


## Log Events: The Atomic Unit of Logging

LogEvents are objects that store all information collected by the Logger. They
are passed on to Appenders that output them, but Appenders usually don't 
utilise all the information present in a log event. The last event produced
by a Logger is stored in its last_event field.

```{r}
yog$info("Vampire stories are generally located in Styria")
yog$last_event  # a summary output of the event
yog$last_event$values  # all values stored in the event as a list
```

LogEvents can contain not only these standard values, but an arbitrary number
of extra values. These extra values are passed as named arguments to the logging
function (as opposed as to paramters to `sprintf()`, which are passed as 
unnamed arguments). It is up to the Appender whether to process them further or 
not. 

You should consider making use of custom fields liberally and using output
formats that support them (such as JSON), rather than producing elaborately 
formatted but hard to parse log messages. If you ever want to analyse your 
log data, you will thank me for this tip.

```{r}
# bad
yog$info("Processing track '%s' with %s waypoints", "track.gpx", 32)

# Good
tf <- tempfile()
yog$add_appender(AppenderJson$new(tf), "json")
yog$info("Processing track", file = "track.gpx", waypoints = 32)
yog$appenders$json$data

```
```{r echo = FALSE}
yog$remove_appender("json")
unlink(tf)
```





## Thresholds & Filters: Controlling Output Detail {#thresholds}

To control the level of detail of the log output, you can set **thresholds** for
Loggers and Appenders. A Logger with a threshold of `warn` will only create
LogEvents of the priorities `warn`, `error` and `fatal` and dispatch them
to its Appenders.

> A **threshold** of a Logger or Appender is the minimum **log level** a 
> LogEvent must have so that that Logger/Appender processes it.

If you require more complex logic to decide whether a LogEvent should be
created/processed you can also assign **filters** to Loggers/Appenders. They
will be applied after the threshold is checked. 

examples:
```{r}
# Only log on weekends
f1 <- function(event, obj) { as.POSIXlt(Sys.Date())$wday %in% c(7, 0) }
# Only log messages that contain the word bird
f2 <- function(event, obj) { grepl("bird", event$msg) }
yog$set_filters(list(f1, f2))

# since this is not very useful, we remove the filters again
yog$set_filters(NULL)

```


## Appenders: Managing Log Destinations

The root logger only logs to the console by default. If you want to redirect 
the output to a file you can just add a file appender to yog.

```{r}
tf <- tempfile()

# Add a new appender to a logger. We don't have to supply a name, but that
# makes it easier to remove later.
yog$add_appender(AppenderFile$new(file = tf), name = "file")

# configure yog so that it logs everything to the file, but only info and above
# to the console
yog$set_threshold(NA)
yog$appenders$console$set_threshold("info")
yog$appenders$file$set_threshold(NA)
yog$info("Another informational message")
yog$debug("A debug message not shown by the console appender")

readLines(tf)

# Remove the appender again
yog$remove_appender("file")
```
```{r echo = FALSE}
unlink(tf)
```


## Inheritance: Hierarchical Loggers {#hierarchy}

Logger hierarchies are a powerful concept to organize logging for different
parts of a larger system. This is mainly relevant for package developers. It
is good practice to have a sepparate Logger for each package. Since it is not
common in R to build complex systems of hierarchically organised packages,
hierarchies will usually be pretty flat (i.e. most Loggers will only inherit
from the root logger). 

Each newly created Logger is child to a parent Logger. If no parent Logger
is specified manually during a Logger's creation, its parent is the Root 
Logger. A logger dispatches the LogRecords it creates not only to its own
Appenders, but also to the Appenders of all its ancestral Loggers (ignoring
the threshold and Filters of the ancestral Loggers, but not of the ancestral 
Appenders). 

If all this sounds confusing to you, take a look at the
[examples](#examples-hierarchy). The common use cases are pretty easy to 
understand and illustrate the *how* and *why* pretty well.



### Log Flow

The graph bellow outlines the flow of LogEvents through the logging system.
This is an important reference if you want to work with Filters and Logger
hierarchies.


![](log_flow.svg)


# Examples {#examples}

## Logging to the console 

yog comes with simple formatting syntax. The configuration possibilites are 
limited, but should be sufficient for most users.

```{r}
lg <- Logger$new(
  "test", 
  appenders = list(cons = AppenderConsole$new()), 
  propagate = FALSE
)

lg$info("the default format")

lg$appenders$cons$layout$set_fmt("%L (%n) [%t] %c(): !! %m !!")
lg$info("A more involved custom format")
```


## Logging to JSON files

JavaScript Object Notation (JSON) is an open-standard file format that uses 
human-readable text to transmit data objects consisting of attribute–value 
pairs and array data types ([wikipedia](https://en.wikipedia.org/wiki/JSON)).
JSON is the **recommended text-based logging format when logging to files**
  ^[Technically, the logger does not produce standard JSON files but 
  [JSON lines](http://jsonlines.org/)], 
as it is human- as well as machine readable. You should only log to a different 
format if you have very good reasons for it. 


```{r}
# install.packages("jsonlite")
tf <- tempfile()

# AppenderJson is just an AppenderFile with LayotJson as default instead 
# and a few extra features
lg <- Logger$new(
  "test logger",
  appenders = AppenderJson$new(file = tf), 
  propagate = FALSE
)

lg$info("JSON naturally supports custom log fields", field = "custom")
lg$info("You can serialize most R data types to JSON", numbers = 1:5)
lg$info("If you ever want to analyse your log files", use = "JSON")

```

JSON is easy to parse and analyse with R
```{r}
read_json_lines(tf)
```

It is also human readable, though maybe this vignette does not 
transport that fact very well because of the lack of horizontal space
```{r}
cat(readLines(tf), sep = "\n\n")
```

```{r, echo = FALSE}
unlink(tf)
```


## Logger Hierarchies  {#examples-hierarchy}

The most common usecases for creating a new Logger rather than just using the
root Logger is if you create a Package that should contain logging. This
way you can have separate Appenders (e.g logfiles) and thresholds for each
package. 

```{r}
# The logger name should be the same as the package name
tf <- tempfile()
lg <- Logger$new("mypackage", appenders = AppenderFile$new(tf))  
```

The `print()` method for Loggers gives a nice overview of the newly created
logger:

```{r}
print(lg)
```

This tells us that `lg` logs all events of at least level `info`. It does 
have a single (unnamed) Appender that logs to a temporary file, and dispatches 
all LogEvents it creates to the Appenders of the root Logger (ignoring the 
threshold and filters of the root Logger, but not of the Appenders).

We can use `lg$fatal()`, `lg$info()`, etc.. to log messages with this Logger:
```{r}
lg$info("A test message for lg")
```

If we do not want `lg` to dispatch to the root Logger, we can set 
`propagate` to `FALSE`.
```{r}
lg$set_propagate(FALSE)
```
When we take a look at the Logger again, we now see that it does not inherit
any Appenders anymore
```{r}
print(lg)
```

Consequently, `lg` no longer outputs log messages to he console
```{r}
lg$info("Nothing to see here")
```
```{r echo = FALSE}
unlink(tf)
```


