---
title: "Yet Another Log4j Inspired Logging Framework"
author: "Stefan Fleck"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{yog}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(yog)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Yog is an [Apache Log4j](https://logging.apache.org/log4j/2.x/) inspired
logging framework based on [R6 classes](https://github.com/r-lib/R6) that 
focuses on extensibility, ease of use, and the ability to use multiple 
appenders with custom thresholds for a single logger.

# Basic usage

Yog comes with a pre-configured logger included. You can just start logging 
with:

```{r}
yog$info("This is an informational message")
yog$fatal("This is an important message about something going wrong")
```

This only logs to the console. If you want to redirect the output to a file
you can just add a file appender to yog

```{r}
tf <- tempfile()
yog$add_appender(AppenderFile$new(file = tf))
readLines(tf)
```

# Structure

* A **Logger** collects information and dispatches it to the appropriate appenders
* An **Appender** writes the log message to destination (the console, a file, 
  a database, etc...)
* A **Layout** is




## Log Levels

Yog supports the standard log4j log levels outlined bellow. Log levels have
numeric and character representations. Creation of custom log levels are not 
yet supported by yog.

```{r, echo = FALSE}

ll <- data.frame(
  `Level` = seq(100, 600, by = 100),
  `Name` = c("FATAL", "ERROR", "WARNING", "INFO", "DEBUG", "TRACE"),
  `Description` = c(
    "Critical error that leads to program abort. Should always indicate a `stop()` or similar",
    "A severe error that does not trigger program abort",
    "A potentially harmful situation, like `warning()`",
    "An informatinal message on the progress of the application",
    "Finer grained informational messages that are mostly useful for debugging",
    "An even finer grained message than debug"
  )
) 

knitr::kable(ll)

```



## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
